---
title: "Data Import"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data 

## Import the FAS Litters CSV

Import and clean the data with janitor package 
```{r}
litters_df <- read_csv("Data_examples/FAS_litters.csv")

litters_df <- janitor::clean_names(litters_df)
```


## Look at the data set 

```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```


```{r eval=FALSE}
view(litters_df)
```


## Importing FAS Pups Dataset

Importing and Cleaning the FAS Pups dataset
```{r}
pups_df <- read_csv("Data_examples/FAS_pups.csv")

pups_df <- janitor::clean_names(pups_df)
```

## Look at read_csv options 

col_names and skipping rows 

```{r}

litters_df <- 
  read_csv(
  file = "Data_examples/FAS_litters.csv", 
  col_names = FALSE, 
  skip = 1
  )
```

what about missing data?

```{r}
litters_df <- 
  read_csv(
    file = "Data_examples/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df <- janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)

mean(pull(litters_df, gd0_weight), na.rm = TRUE)
```

What if we code `group` as as factor variable? 

```{r}
litters_df <- 
  read_csv(
    file = "Data_examples/FAS_litters.csv",
    na = c("NA", "", "."), 
    col_types = cols(
      Group = col_factor()
    )
  )
litters_df
```

## Import an excel file

Import MLB 2011 summary data

```{r}
mlb_df <- 
  read_excel("Data_examples/mlb11.xlsx", sheet = "mlb11")
```

## Import SAS data 

```{r}
pulse_df <- read_sas("Data_examples/public_pulse_data.sas7bdat")
```


## Never use read.csv()

```{r}
litters_df <- read.csv("Data_examples/FAS_litters.csv")

```

Never do this either: 
```{r}
litters_df$L
```

